---
title Guide to Regression Discontinuity Design
output
  html_document
    df_print paged
editor_options
  chunk_output_type inline
---

# Analysis of Regression Discontinuity Design

This code uses simulated data based on a study of professional development of early childhood educators in Florida.
Teacherstudent interactions in early childhood classroom were evaluated with the Classroom Assessment Scoring system (CLASS)
Teachers of classrooms scoring below a 3 had to enroll in professional development
The CLASS was also used as the outcome, measured at the end of the year


# Read and check the dataset
load(file="U:/U-Shu-Ren/Mine/Data_Certificate/DSMDLP Training Fall 2024/Module_13/simulated_RDD_data.Rdata")

#Read and check the dataset
```{r}
load(file="U:/U-Shu-Ren/Mine/Data_Certificate/DSMDLP Training Fall 2024/Module_13/simulated_RDD_data.Rdata")
summary(data)
```
```{r}
names(data)
```
Plot the observations and regression lines

```{r}

# Plot the data
with(data, plot(CLASS.post ~ CLASS.pre, col=NULL, xlab="CLASS.pre",
     ylab="CLASS.post", ylim=c(min(CLASS.post)-10, max(CLASS.post)+10)))

# Obtain index of those treated
tx <- which(data$PD == 1)  # obtain index of those treated

# Add points to the plot
with(data, points(CLASS.pre[tx], CLASS.post[tx], col="green", pch=4))
with(data, points(CLASS.pre[-tx], CLASS.post[-tx], col="red",pch=5))


```

=====================================================
Center assignment variable
```{r}
cutoff=3
data$CLASS.pre = data$CLASS.pre - cutoff
summary(data$CLASS.pre)
```

# Estimate Local Average Treatment Effect with the ANCOVA model

```{r}


# Fit the linear model
model <- lm(CLASS.post ~ PD + CLASS.pre + I(CLASS.pre^2), data)

model = lm(CLASS.post ~ PD + CLASS.pre + I(CLASS.pre^2), data)
summary(model)
```



Add the fitted lines to plot
```{r}
with(data,plot(CLASS.post ~ CLASS.pre, col=NULL, xlab=CLASS.pre,
     ylab=CLASS.post, ylim=c(min(CLASS.post)-10, max(CLASS.post)+10)))

tx - which(data$PD==1)  #obtain index of those treated.
with(data, points(CLASS.pre[tx], CLASS.post[tx], col="cyan", pch=4))
with(data,points(CLASS.pre[-tx], CLASS.post[-tx], col="yellow"))
coefs <- coefficients(model)

pred.control <- function(pre) {
  coefs[1] + coefs[3] * pre + coefs[4] * pre^2
}
pred.control = function(pre) {
  coefs[1] + coefs[3]*pre + coefs[4]*pre^2}
pred.treat = function(pre) {coefs[1] + coefs[2] + coefs[3]*pre + coefs[4]*pre^2}
with(data,curve(pred.control, from=min(CLASS.pre), to=0, add=T, col="blue"))
with(data,curve(pred.treat, 0, max(CLASS.pre)+1, add=T, col="red"))
```

================================================

# Estimate the local average treatment effect with local linear regression

Use treatment observations to fit a local linear regression
```{r}
lo.tx <- loess(CLASS.post ~ CLASS.pre, data[data$PD==1,],span=0.75,family="gaussian", surface="direct")
summary(lo.tx)
```


Use control observations to fit a local linear regression
```{r}
lo.c <- loess(CLASS.post ~ CLASS.pre, data[data$PD==0,], span=0.75,family="gaussian", surface="direct")
summary(lo.c)
```

Predict the expected value of the treated at the cutoff
```{r}
yhat.tx <- predict(lo.tx, newdata=0, se=T) #predict value of treated at the cutoff
yhat.tx
```

Predict the expected value of the control at the cutoff
```{r}
yhat.c <- predict(lo.c, newdata = 0, se=T) #predict value of untreated at the cutoff
yhat.c
```


Estimate the local average treatement effect
```{r}
effect = yhat.tx$fit - yhat.c$fit  #estimate treatment effect
effect
```


calculate confidence interval of the LATE
```{r}
lower.limit = effect - 1.96*(yhat.tx$se.fit + yhat.c$se.fit)
upper.limit = effect + 1.96*(yhat.tx$se.fit + yhat.c$se.fit)
paste(lower.limit,upper.limit)
```
